Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex
second pass on CG

\textcolor{red}{Again, should we highlight the differences between these methods and ours? I prefer to do but wouldn't it sounds bad?}

@@

\subsubsection{Fast Whitening}

vector $x$ requires a system of linear equations $\Sigma w = (x - \mu)$ to be
solved, usually by decomposing the covariance matrix $\Sigma$ (with LU or
Cholesky decomposition). Since the covariance matrices of HOG templates have
nice spectral properties (see Section \ref{sec:fast_whitening}), the conjugate
gradient method will converge to the solution several orders of magnitude
faster than decomposition methods.

in the background will be incorporated into the discriminative template.
Alternatively, we could leave the background textureless \scream{did Hariharan
do this?} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG
template produces in undesirable background artifacts. These are created when
centering the template (by subtracting the mean $\mu$), where strong negative
weights are introduced in the textureless region (as seen in
Fig.~\ref{fig:whocomparison}). This could result in positive matches being
suppressed due to spurious background gradients.

gradients. After whitening we restore the full template, with the zero-gradient
cells left unchanged. This approach has the added benefit of reducing the size
of the linear system we need to solve, further speeding up our method.

NZ-WHO templates, illustrating that our method is faster and more
discriminative than Whitened Histograms of Orientations. \scream{Does the table
show this? should we do more than AP and realtime analysis?}
% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

We use the GPU to parallelize the lookup of the correct autocovariance cells
during covariance matrix syntheses, which is an order of magnitude times faster than using
the CPU (Figure \ref{fig:covariancetime}).

@@

\subsection{Fast Whitening and Convolution}
\label{sec:fast_whitening}
\begin{figure}[t]
  \begin{center}
     \includegraphics[width=0.9\linewidth]{whiten_orig_depth} 
  \end{center}
  \caption{Realistic rendering and depth}
  \label{fig:rendering}
  \end{figure}

To whiten a signal $x$, we have to solve $\Sigma w = x - \mu$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are Hermitian and positive semidefinite to solve the system via the Cholesky
decomposition, which is twice as fast as the LU decomposition. These
decompositions require $O(n^3)$ time, whereas the conjugate gradient method
takes $O(n^2\kappa)$ time, where $\kappa$ is the condition number of the matrix.
This makes conjugate gradient a faster option for matrices with small condition
numbers relative to their size.

% Figure \ref{fig:spectral} shows the singular values of a covariance matrix for
% one HOG template after a small regularization constant is added to the diagonal.
% The condition number of of the matrix is , much smaller than the size
% of the matrix (7500). As a result, conjugate gradient converges in 50
% ms, two orders of magnitude faster than the system can be solved using the
% Cholesky factorization (5 sec).

% The final challenge in on-the-fly template evaluation is performing the
% convolution between the template and the target image. Standard convolution
% would be prohibitively slow for the high resolution templates we're using, so we
% implemented the convolution via Fast Fourier Transform on the GPU. 
% 
% This allows
% template evaluation to run in \scream{x ms}, compared to \scream{x ms} for the
% naive CPU implementation.

@@

\subsection{Variations of WHO}

We empirically found out that NZ-WHO performs reasonably well without time
consuming calibration stage compare to other variational method to generate
templates. We used 1 CAD model with viewpoints covering 24 azimuths and 4
elevations. Each template takes approximately 80 milliseconds to generate a
NZ-WHO template. We also calibrated templates using the method presented in
\cite{Aubry14}. The calibration learns affine transformation of the detection
confidence.% Note that in Section \ref{experiments}, we provide the same
experiment with 9 CAD models.

@@

\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. Since naive convolution takes
$O(nm)$ time where $n$ is the size of the signal and $m$ is the size of the template.
Whereas, FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. 
For large $m$, high resolution template, we can gain computational advantage.

% We pad zero to the 2D image features by the largest kernel size. And we
% transform the kernel to match the padded image feature. To compute convolution,
% we compute element-wise product with the transformed kernel and sum all the
% features. Unlike \cite{Podlozhnyuk}, we do not need to shift the data nor
% kernel if we use simple mathematical trick. 
%   
% \begin{align}
%     x_{pad} & = \mathcal{P}_{n+m}(x)\\
%     y_{pad} & = \mathcal{P}_{n+m}(y)\\
%     x_{pad} \ast y_{pad} & = \mathcal{F}^{-1}(\mathcal{F}(x_{pad}) \circ \mathcal{F}(y_{pad}))
% \end{align}
% 
% Where $\ast$ is the circular convolution and $\circ$ is the Hadamard Product
% and $\mathcal{F}$ is the Fourier Transformation and $\mathcal{P}_n$ is the
% padding operation that append zeros to make vector of size $n$.
% 
% For each dimension of features, we transform kernel and image feature into
% frequency domain using Fast Fourier Transformation. We transform the image into
% frequency domain once and use the transformed image for all templates.
% 
% We implemented the convolution using FFT on GPU and got 10x speed up.

% [Figure for convolution time for different template resolution, CPU, GPU, naive convolution]
%------------------------------------------------------------------------

\section{Continuous Viewpoint Estimation using MCMC}
\label{sec:fine}
% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation
and evaluation computationally inexpensive. This means we can efficiently
explore the parameter space to find the best object pose, scale, model type and
camera focal length.

@@

Suppose that $x = [p, m, f]$ where $p$ is the 3D rotation of the CAD model and
$m$ is the CAD model index and $f$ is the focal length. Then we sample $x$ using 
% there are several advantages that was not possible such as \textbf{on-the-fly} template generation and making large number of templates  
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Different modes of Metropolis Hasting proposals, see Section \ref{sec:fine}}
 \label{fig:tuningmode}
\end{figure}

>>>>

\textcolor{red}{Again, should we highlight the differences between these methods and ours?}

@@

\subsubsection{Fast Whitening using Conjugate Gradient}

The computational efficiency of our approach comes from using the iterative conjugate gradient method to whiten the HOG template. Whitening the non-zero feature vector $x'$ requires a system of linear equations $\Sigma' w' = (x' - \mu')$ to be solved. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition, which requires $O(n^3)$ time.

The conjugate gradient method is an iterative algorithm for solving symmetric positive definite systems which runs in $O(n^2\kappa)$ time, where $\kappa$ is the condition number of the matrix.
This makes conjugate gradient faster than decomposition for matrices with small condition
numbers relative to their size.

The covariance matrix for HOG templates is typically ill-conditioned, but we can add a regularization constant to the diagonal to reduce its condition number to the point where the conjugate gradient method will converge quickly. We use a constant of $0.15$, which reduces the condition number from $10^{20}$ to $50$, much smaller than the dimension of the matrix (7000).

As a result, a GPU implementation of conjugate gradient converges in \scream{x
ms}, two orders of magnitude faster than the system can be solved using the
Cholesky factorization (\scream{5 s}).

\subsubsection{Fast Convolution using the FFT}

% We compare the performance of WHO and NZ-WHO on 3D Object dataset \cite{Savarese07} and report performance on Table \ref{tab:who_initializations}.

@@

We use the GPU to parallelize the lookup of the correct autocovariance cells
during covariance matrix syntheses, which is \scream{x} times faster than using
the CPU (Figure \ref{fig:covariancetime}).

@@

The final challenge in on-the-fly template evaluation is performing the
convolution between the template and the target image. Standard convolution
would be prohibitively slow for the high resolution templates we're using, so we
implemented the convolution via Fast Fourier Transform on the GPU. This allows
template evaluation to run in \scream{x ms}, compared to \scream{x ms} for the
naive CPU implementation.

@@

\subsection{Variations of WHO}
\scream{sam thinks this should be moved to the experiments section}
We empirically found out that NZ-WHO performs reasonably well without time
consuming calibration stage compare to other variational method to generate
templates. We used 1 CAD model with viewpoints covering 24 azimuths and 4
elevations. Each template takes approximately 80 milliseconds to generate a
NZ-WHO template. We also calibrated templates using the method presented in
\cite{Aubry14}. The calibration learns affine transformation of the detection
confidence.% Note that in Section \ref{experiments}, we provide the same
experiment with 9 CAD models.

@@

\begin{comment}
\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture details of an object to give accurate 2D-3D matching. These large templates cause computational burden when computing convolution. Though good for detecting accurate model and pose, convolution of high-resolution templates are much more slower since computation time scales linearly with the number of HOG cells in the template. To overcome and speed up the convolution, we propose FFT based GPU convolution which scales. Suppose there are length $n$ signal and length $m$ filter, naive convolution takes $O(nm)$ time where as FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large $m$, high resolution template, we can gain computational advantage.

we pad zero to the 2D image features by the largest kernel size. And we transform the kernel to match the padded image feature. To compute convolution, we compute element-wise product with the transformed kernel and sum all the features. Unlike \cite{Podlozhnyuk}, we do not need to shift the data nor kernel if we use simple mathematical trick.

\begin{align}
    x_{pad} & = \mathcal{P}_{n+m}(x)\\
    y_{pad} & = \mathcal{P}_{n+m}(y)\\
    x_{pad} \ast y_{pad} & = \mathcal{F}^{-1}(\mathcal{F}(x_{pad}) \circ \mathcal{F}(y_{pad}))
\end{align}

  Where $\ast$ is the circular convolution and $\circ$ is the Hadamard Product and $\mathcal{F}$ is the Fourier Transformation and $\mathcal{P}_n$ is the padding operation that append zeros to make vector of size $n$.

For each dimension of features, we transform kernel and image feature into frequency domain using Fast Fourier Transformation. We transform the image into frequency domain once and use the transformed image for all templates.

We implemented the convolution using FFT on GPU and got 10x speed up.

[Figure for convolution time for different template resolution, CPU, GPU, naive convolution]
\end{comment}
%------------------------------------------------------------------------

\section{Estimating Parameters using MCMC}
\label{sec:fine}
% Estimating viewpoint continuously to fine tune a initial matching
The NZ-WHO template matching method we have presented makes template generation
and evaluation computationally inexpensive. This means we can efficiently
explore the parameter space to find the best object pose, scale, model type and
camera focal length.

Our method begins by convolving the

@@

% there are several advantages that was not possible such as \textbf{on-the-fly} template generation and making large number of templates  
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Different modes of Metropolis Hasting proposals, see Section \ref{sec:fine}}
 \label{fig:tuningmode}
\end{figure}
